---
title: "Bluefish wind analysis"
author: "Abigail Tyrell"
date: "`r format(Sys.time(), '%d %b %Y')`"
output:
  html_document:
    toc: true
    number_sections: true
    toc_depth: 4
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
`%>%` <- magrittr::`%>%`
```

# Introduction

This wind analysis was inspired by Munch and Conover (2000), who found a correlation between wind speed and coastwide YOY index. The wind measurements used are described below: 

> The NOAA environmental buoy array provided
environmental data from 1970 to 1992. Buoys collected
hourly records of windspeed and wind direction at a
height of 10 m, and surface temperature at a depth of
1.5 m. Data were examined from a total of 34 buoys in
fixed locations throughout the region from Cape Fear to
the Gulf of Maine and extending from nearshore
to beyond the shelf break. Buoy locations are reported
in Munch (1997).

Spring-spawned YOY were apparently characterized as fish between 13-27cm in length in the autumn bottom trawl, but "A complete set of length-frequency distributions and cohort designations is presented in Munch (1997)".

And the observed correlation is described:

> A strong relationship existed between offshore wind
and recruitment. Using north (n) and east (e) windstress components from NOAA buoys near the
edge of the shelf of the Mid-Atlantic Bight, the
model ln(c.p.u.e.) = -16.5 n - 12.8 e + 0.59 (p<0.05)
explained 36% of the variance in spring cohort abundance in the SNE region

I am trying to obtain a copy of Munch (1997), which is a master's thesis (Munch, S. B. 1997. Recruitment Dynamics of Bluefish, Pomatonus saltatrix, on the continental shelf from Cape Fear to
Cape Cod, 1973–1995. M.S. Thesis. State University of New York at Stony Brook).

# Determine relevant strata

## What polygons to use?

Entire east coast:

```{r}
# setup
xmin <- -80
xmax <- -65
ymin <- 30
ymax <- 45
xlims <- c(xmin, xmax)
ylims <- c(ymin, ymax)
crs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

ne_countries <- rnaturalearth::ne_countries(
  scale = 10,
  continent = "North America",
  returnclass = "sf"
) %>%
  sf::st_transform()

ne_states <- rnaturalearth::ne_states(
  country = "united states of america",
  returnclass = "sf"
) %>% sf::st_transform()


# strata
strata <- NEesp::latlong %>%
  dplyr::select(strata) %>%
  dplyr::distinct() %>%
  dplyr::rename(STRATA = strata)

new_shape <- NEesp::shape %>%
  dplyr::select(STRATA, geometry) %>%
  sf::st_transform(proj4string = crs)
#sf::st_crs(new_shape) <- crs

strata_plot <- dplyr::full_join(new_shape, strata)

# plot
p1 <- ggplot2::ggplot() +
  ggplot2::geom_sf(
    data = new_shape, # all survey strata
    # fill = "aliceblue",
    ggplot2::aes(fill = STRATA),
    alpha = 0.9, 
    size = 0.01, 
    color = "grey30"
  ) +
  # ggplot2::geom_sf(data = strata_plot,
  #     size = 0.05, color = "grey30") +
  ggplot2::geom_sf(
    data = ne_countries,
    color = "grey60", 
    size = 0.25
  ) +
  ggplot2::geom_sf(
    data = ne_states,
    color = "grey60", 
    size = 0.05
  ) +
  viridis::scale_fill_viridis(discrete = FALSE) +
  ggplot2::coord_sf(crs = crs, 
                    xlim = xlims, 
                    ylim = ylims) +
  ggthemes::theme_map() +
  ggplot2::theme(legend.direction = "horizontal")
print(p1)
```

Inshore (ish) strata:

```{r}
p1 <- ggplot2::ggplot() +
  ggplot2::geom_sf(
    data = new_shape, # all survey strata
    fill = "aliceblue",
    #  ggplot2::aes(fill = STRATA),
    alpha = 0.9, 
    size = 0.01, 
    color = "grey30"
  ) +
  ggplot2::geom_sf(
    data = new_shape %>%
      dplyr::filter(STRATA > 2000 & STRATA < 8000),
    fill = "maroon",
    size = 0.05, 
    color = "grey30"
  ) +
  ggplot2::geom_sf(
    data = ne_countries,
    color = "grey60",
    size = 0.25
  ) +
  ggplot2::geom_sf(
    data = ne_states,
    color = "grey60", 
    size = 0.05
  ) +
  viridis::scale_fill_viridis(discrete = FALSE) +
  ggplot2::coord_sf(crs = crs, 
                    xlim = xlims, 
                    ylim = ylims) +
  ggthemes::theme_map()
print(p1)
```

All strata south of ~ Cape Cod

```{r}

large_geom <- new_shape %>% 
    dplyr::summarise(geometry = sf::st_union(geometry)) %>%
  sf::st_crop(y = c(xmin = -80, xmax = -69, 
                    ymax = 41.5, ymin = 32))

p1 <- ggplot2::ggplot() +
  ggplot2::geom_sf(
    data = new_shape, # all survey strata
    fill = "aliceblue",
    #  ggplot2::aes(fill = STRATA),
    alpha = 0.9, 
    size = 0.01, 
    color = "grey30"
  ) +
  ggplot2::geom_sf(
    data = large_geom,
    fill = "maroon",
    size = 0.05, 
    color = "grey30"
  ) +
  ggplot2::geom_sf(
    data = ne_countries,
    color = "grey60",
    size = 0.25
  ) +
  ggplot2::geom_sf(
    data = ne_states,
    color = "grey60", 
    size = 0.05
  ) +
  viridis::scale_fill_viridis(discrete = FALSE) +
  ggplot2::coord_sf(crs = crs, 
                    xlim = xlims, 
                    ylim = ylims) +
  ggthemes::theme_map()
print(p1)
```

# Daily data

## Downloaded daily mean dataset (uwind, vwind) from NOAA PSL
Located in `data-raw` folder.

## Extract mean wind by day

### v wind

#### Read in wind data
```{r, echo = TRUE}

# use ecopull function
devtools::install_github("kimberly-bastille/ecopull")
vwind <- ecopull::nc_to_raster(nc = here::here("data-raw/vwnd.10m.2021.nc"),
                               varname = "vwnd",
                               show_images = TRUE) 

vwind <- raster::rotate(vwind)
raster::plot(vwind)
```

#### Mask to relevant strata
```{r, echo = TRUE}
m_vwind <- raster::mask(x = vwind, 
                        mask = new_shape[new_shape$STRATA > 2000 & 
                                           new_shape$STRATA < 8000,2])
mean(m_vwind@data@values, na.rm = TRUE)
m_vwind[[1]]

raster::plot(m_vwind)
```

##### Plot over map
```{r}
rast_df <- raster::as.data.frame(m_vwind, xy = TRUE)

p1 +
  ggplot2::geom_tile(data = rast_df[,1:3] %>%
                       tidyr::drop_na(),
                     ggplot2::aes(x = x,
                                  y = y,
                                  fill = X2021.01.01))
```

Seems like the existing strata are too small, try masking to more strata.

#### Mask to different strata

```{r, echo = TRUE}
m_vwind2 <- raster::mask(x = vwind, 
                        mask = large_geom)
mean(m_vwind2@data@values, na.rm = TRUE)
m_vwind2[[1]]

raster::plot(m_vwind2)
```

##### Plot over map
```{r}
rast_df <- raster::as.data.frame(m_vwind2, xy = TRUE)

p1 +
  ggplot2::geom_tile(data = rast_df[,1:3] %>%
                       tidyr::drop_na(),
                     ggplot2::aes(x = x,
                                  y = y,
                                  fill = X2021.01.01)) +
  ggplot2::theme(legend.direction = "horizontal")
```

This seems like a better geometry.

### u wind

#### Read in wind data
```{r, echo = TRUE}

# use ecopull function
devtools::install_github("kimberly-bastille/ecopull")
uwind <- ecopull::nc_to_raster(nc = here::here("data-raw/uwnd.10m.2021.nc"),
                               varname = "uwnd",
                               show_images = TRUE) 

uwind <- raster::rotate(uwind)
raster::plot(uwind)
```

#### Mask to strata

```{r, echo = TRUE}
m_uwind2 <- raster::mask(x = uwind, 
                        mask = large_geom)
mean(m_uwind2@data@values, na.rm = TRUE)
m_uwind2[[1]]

raster::plot(m_uwind2)
```

##### Plot over map
```{r}
rast_df <- raster::as.data.frame(m_uwind2, xy = TRUE)

p1 +
  ggplot2::geom_tile(data = rast_df[,1:3] %>%
                       tidyr::drop_na(),
                     ggplot2::aes(x = x,
                                  y = y,
                                  fill = X2021.01.01)) +
  ggplot2::theme(legend.direction = "horizontal")
```

### Calculate weekly mean u and v wind

#### v wind 

Break into north and south regions at 39N (approximately the bottom of NJ)

Positive v wind = wind is blowing to the north

```{r}
north <- large_geom %>%
  sf::st_crop(y = c(xmin = -80, xmax = -69, 
                    ymax = 41.5, ymin = 39.001))

south <- large_geom %>%
  sf::st_crop(y = c(xmin = -80, xmax = -69, 
                    ymax = 39, ymin = 32))

north_vwind <- raster::mask(x = vwind, 
                        mask = north)

south_vwind <- raster::mask(x = vwind, 
                        mask = south)

"names" <- NULL
"values" <- NULL
df<- data.frame()
for(i in 1:raster::nlayers(north_vwind)){
  names <- north_vwind[[i]]@data@names
  mean_wind <- mean(north_vwind[[i]]@data@values, na.rm = TRUE)
  min_wind <- north_vwind[[i]]@data@min
  max_wind <- north_vwind[[i]]@data@max
  
  dat<-cbind(names, mean_wind, min_wind, max_wind)
  df<-rbind(df, dat)
  }
head(df)

north_df <- df %>%
  dplyr::mutate(Year = stringr::str_extract(names, pattern = "\\d{4}"), 
                                names = stringr::str_remove(names, pattern = "X"), 
                                DOY = lubridate::as_date(names),
                week = lubridate::week(DOY),
                mean_wind = as.numeric(mean_wind),
                min_wind = as.numeric(min_wind),
                max_wind = as.numeric(max_wind))

# south region

"names" <- NULL
"values" <- NULL
df<- data.frame()
for(i in 1:raster::nlayers(south_vwind)){
  names <- south_vwind[[i]]@data@names
  mean_wind <- mean(south_vwind[[i]]@data@values, na.rm = TRUE)
  min_wind <- south_vwind[[i]]@data@min
  max_wind <- south_vwind[[i]]@data@max
  
  dat<-cbind(names, mean_wind, min_wind, max_wind)
  df<-rbind(df, dat)
  }
head(df)

south_df <- df %>%
  dplyr::mutate(Year = stringr::str_extract(names, pattern = "\\d{4}"), 
                                names = stringr::str_remove(names, pattern = "X"), 
                                DOY = lubridate::as_date(names),
                week = lubridate::week(DOY),
                mean_wind = as.numeric(mean_wind),
                min_wind = as.numeric(min_wind),
                max_wind = as.numeric(max_wind))

daily_wind <- rbind(north_df %>% 
                      dplyr::mutate(region = "North"),
                    south_df %>% 
                      dplyr::mutate(region = "South"))

ggplot2::ggplot(daily_wind,
                ggplot2::aes(x = DOY,
                             y = mean_wind %>%
                               as.numeric(),
                             color = region)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme_bw() +
  ggplot2::ggtitle("Daily v wind")

v_weekly <- daily_wind %>%
  dplyr::group_by(week, region) %>%
  dplyr::summarise(mean_wind_wk = mean(mean_wind %>% as.numeric(), na.rm = TRUE),
                   mean_min = mean(min_wind %>% as.numeric(), na.rm = TRUE),
                   mean_max = mean(max_wind %>% as.numeric(), na.rm = TRUE)) 

v_weekly %>%
  ggplot2::ggplot(ggplot2::aes(x = week,
                             y = mean_wind_wk,
                             color = region)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme_bw() +
  ggplot2::ggtitle("Weekly v wind")
```

#### u wind

Positive u wind = wind is blowing to the east

```{r}
north_uwind <- raster::mask(x = uwind, 
                        mask = north)

south_uwind <- raster::mask(x = uwind, 
                        mask = south)

"names" <- NULL
"values" <- NULL
df<- data.frame()
for(i in 1:raster::nlayers(north_uwind)){
  names <- north_uwind[[i]]@data@names
  mean_wind <- mean(north_uwind[[i]]@data@values, na.rm = TRUE)
  min_wind <- north_uwind[[i]]@data@min
  max_wind <- north_uwind[[i]]@data@max
  
  dat<-cbind(names, mean_wind, min_wind, max_wind)
  df<-rbind(df, dat)
  }
head(df)

north_df <- df %>%
  dplyr::mutate(Year = stringr::str_extract(names, pattern = "\\d{4}"), 
                                names = stringr::str_remove(names, pattern = "X"), 
                                DOY = lubridate::as_date(names),
                week = lubridate::week(DOY))

# south region

"names" <- NULL
"values" <- NULL
df<- data.frame()
for(i in 1:raster::nlayers(south_uwind)){
  names <- south_uwind[[i]]@data@names
  mean_wind <- mean(south_uwind[[i]]@data@values, na.rm = TRUE)
  min_wind <- south_uwind[[i]]@data@min
  max_wind <- south_uwind[[i]]@data@max
  
  dat<-cbind(names, mean_wind, min_wind, max_wind)
  df<-rbind(df, dat)
  }
head(df)

south_df <- df %>%
  dplyr::mutate(Year = stringr::str_extract(names, pattern = "\\d{4}"), 
                                names = stringr::str_remove(names, pattern = "X"), 
                                DOY = lubridate::as_date(names),
                week = lubridate::week(DOY))

# plots

daily_wind <- rbind(north_df %>% 
                      dplyr::mutate(region = "North"),
                    south_df %>% 
                      dplyr::mutate(region = "South"))

ggplot2::ggplot(daily_wind,
                ggplot2::aes(x = DOY,
                             y = mean_wind %>%
                               as.numeric(),
                             color = region)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme_bw() +
  ggplot2::ggtitle("Daily u wind")

u_weekly <- daily_wind %>%
  dplyr::group_by(week, region) %>%
  dplyr::summarise(mean_wind_wk = mean(mean_wind %>% as.numeric(), na.rm = TRUE),
                   mean_min = mean(min_wind %>% as.numeric(), na.rm = TRUE),
                   mean_max = mean(max_wind %>% as.numeric(), na.rm = TRUE)) 

u_weekly %>%
  ggplot2::ggplot(ggplot2::aes(x = week,
                             y = mean_wind_wk,
                             color = region)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme_bw() +
  ggplot2::ggtitle("Weekly u wind")
```

## Calculate hypotenuse and angle by week

Angles/directions are probably not reliable yet - have to ground truth

```{r}
colnames(v_weekly)[3:5] <- c("v_mean", "v_min", "v_max")
colnames(u_weekly)[3:5] <- c("u_mean", "u_min", "u_max")

total_wind <- dplyr::full_join(v_weekly, u_weekly) %>%
  dplyr::mutate(v_mean = v_mean %>% round(digits = 3),
                u_mean = u_mean %>% round(digits = 3),
                wind_magnitude = sqrt(v_mean^2 + u_mean^2) %>% round(digits = 3),
                
                ns = ifelse(v_mean > 0, "north", "south"),
                ew = ifelse(u_mean > 0, "east", "west"),
                wind_blowing_to = paste0(ns, ew),
                
                angle = atan(v_mean/u_mean) %>% round(digits = 3),
                coast_angle = (pi/4) %>% round(digits = 3), # dummy value
                new_angle = (pi/2 - angle - coast_angle) %>% round(digits = 3),
                
                # have to add sign back somehow
                longshore_wind = (wind_magnitude / sqrt(tan(new_angle)^2 + 1))
                 %>% round(digits = 3),
                crossshore_wind = sqrt(wind_magnitude^2 - longshore_wind^2) 
                %>% round(digits = 3),
                
                along = ifelse(new_angle > 0, "up_coast", "down_coast"),
                across = ifelse(abs(new_angle) < pi/2, "away_coast", "towards_coast"),
                wind_blowing_to2 = paste(along, across)
                ) %>%
  dplyr::select(-c(ns, ew, along, across, v_min, v_max, u_min, u_max))

knitr::kable(total_wind)
```

### Try plotting to see if things look correct

```{r}
ggplot2::ggplot(total_wind %>%
                  dplyr::filter(region == "North"),
                ggplot2::aes(x = 0,
                             y = 0,
                             xend = v_mean,
                             yend = u_mean,
                             color = as.factor(week))) +
  ggplot2::geom_segment() +
# ggplot2::geom_point() +
 # ggplot2::coord_polar() +
  ggplot2::theme_bw()
```
## Calculate cross shore and long shore winds

# Monthly data

## Extract mean wind by month

### v wind

#### Read in wind data
```{r, echo = TRUE}
# use ecopull function
vwind <- ecopull::nc_to_raster(nc = here::here("data-raw/vwnd.10m.mon.mean.nc"),
                               varname = "vwnd",
                               show_images = TRUE) 

vwind <- raster::rotate(vwind)
raster::plot(vwind)
```

#### Mask to relevant strata
```{r, echo = TRUE}
m_vwind <- raster::mask(x = vwind, 
                        mask = large_geom)
mean(m_vwind@data@values, na.rm = TRUE)
m_vwind[[1]]

raster::plot(m_vwind)
```

##### Plot over map
```{r}
rast_df <- raster::as.data.frame(m_vwind, xy = TRUE)

p1 +
  ggplot2::geom_tile(data = rast_df[,1:3] %>%
                       tidyr::drop_na(),
                     ggplot2::aes(x = x,
                                  y = y,
                                  fill = X1979.01.01))
```

### u wind

#### Read in wind data
```{r, echo = TRUE}
# use ecopull function
uwind <- ecopull::nc_to_raster(nc = here::here("data-raw/uwnd.10m.mon.mean.nc"),
                               varname = "uwnd",
                               show_images = TRUE) 

uwind <- raster::rotate(uwind)
raster::plot(uwind)
```

#### Mask to relevant strata
```{r, echo = TRUE}
m_uwind <- raster::mask(x = uwind, 
                        mask = large_geom)
mean(m_uwind@data@values, na.rm = TRUE)
m_uwind[[1]]

raster::plot(m_uwind)
```

##### Plot over map
```{r}
rast_df <- raster::as.data.frame(m_uwind, xy = TRUE)

p1 +
  ggplot2::geom_tile(data = rast_df[,1:3] %>%
                       tidyr::drop_na(),
                     ggplot2::aes(x = x,
                                  y = y,
                                  fill = X1979.01.01))
```

### Calculate monthly mean u and v wind

#### v wind 

Break into north and south regions at 39N (approximately the bottom of NJ)

Positive v wind = wind is blowing to the north

```{r, fig.height = 12}
north_vwind <- raster::mask(x = vwind, 
                        mask = north)

south_vwind <- raster::mask(x = vwind, 
                        mask = south)

"names" <- NULL
"values" <- NULL
df<- data.frame()
for(i in 1:raster::nlayers(north_vwind)){
  names <- north_vwind[[i]]@data@names
  mean_wind <- mean(north_vwind[[i]]@data@values, na.rm = TRUE)
  min_wind <- north_vwind[[i]]@data@min
  max_wind <- north_vwind[[i]]@data@max
  
  dat<-cbind(names, mean_wind, min_wind, max_wind)
  df<-rbind(df, dat)
  }
head(df)

north_df <- df %>%
  dplyr::mutate(Year = stringr::str_extract(names, pattern = "\\d{4}"), 
                                names = stringr::str_remove(names, pattern = "X"), 
                                DOY = lubridate::as_date(names),
                week = lubridate::week(DOY),
                mean_wind = as.numeric(mean_wind),
                min_wind = as.numeric(min_wind),
                max_wind = as.numeric(max_wind))

# south region

"names" <- NULL
"values" <- NULL
df<- data.frame()
for(i in 1:raster::nlayers(south_vwind)){
  names <- south_vwind[[i]]@data@names
  mean_wind <- mean(south_vwind[[i]]@data@values, na.rm = TRUE)
  min_wind <- south_vwind[[i]]@data@min
  max_wind <- south_vwind[[i]]@data@max
  
  dat<-cbind(names, mean_wind, min_wind, max_wind)
  df<-rbind(df, dat)
  }
head(df)

south_df <- df %>%
  dplyr::mutate(Year = stringr::str_extract(names, pattern = "\\d{4}"), 
                                names = stringr::str_remove(names, pattern = "X"), 
                                DOY = lubridate::as_date(names),
                week = lubridate::week(DOY),
                mean_wind = as.numeric(mean_wind),
                min_wind = as.numeric(min_wind),
                max_wind = as.numeric(max_wind))

monthly_wind <- rbind(north_df %>% 
                      dplyr::mutate(region = "North"),
                    south_df %>% 
                      dplyr::mutate(region = "South")) %>%
  dplyr::mutate(month = lubridate::month(DOY),
                year = lubridate::year(DOY))

v_monthly_wind <- monthly_wind

ggplot2::ggplot(monthly_wind,
                ggplot2::aes(x = month,
                             y = mean_wind %>%
                               as.numeric(),
                            # shape = region,
                             color = as.factor(year)
                             )) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme_bw() +
  ggplot2::facet_wrap(~region,
                      ncol = 1) +
  #viridis::scale_color_viridis() +
  ggplot2::ggtitle("Monthly v wind") +
  ggplot2::theme(legend.position = "bottom")
```

##### Test correlation with recruitment index from the 2019 assessment model
```{r, fig.height = 16}
recruit <- NEesp::asmt %>% 
  dplyr::filter(Species == "Bluefish", 
                Metric == "Recruitment", AssessmentYear == 2019) %>%
  dplyr::select(Year, Value) %>%
  dplyr::rename(Recruitment = Value) %>%
  dplyr::mutate(Year = as.character(Year))

dplyr::full_join(recruit, monthly_wind) %>%
  tidyr::drop_na(Recruitment, mean_wind) %>%
  ggplot2::ggplot(ggplot2::aes(x = mean_wind,
                               y = Recruitment)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::facet_grid(cols = ggplot2::vars(region),
                      rows = ggplot2::vars(month)) +
  ggplot2::theme_bw()
```

#### u wind

Positive u wind = wind is blowing to the east

```{r, fig.height = 12}
north_uwind <- raster::mask(x = uwind, 
                        mask = north)

south_uwind <- raster::mask(x = uwind, 
                        mask = south)

"names" <- NULL
"values" <- NULL
df<- data.frame()
for(i in 1:raster::nlayers(north_uwind)){
  names <- north_uwind[[i]]@data@names
  mean_wind <- mean(north_uwind[[i]]@data@values, na.rm = TRUE)
  min_wind <- north_uwind[[i]]@data@min
  max_wind <- north_uwind[[i]]@data@max
  
  dat<-cbind(names, mean_wind, min_wind, max_wind)
  df<-rbind(df, dat)
  }
head(df)

north_df <- df %>%
  dplyr::mutate(Year = stringr::str_extract(names, pattern = "\\d{4}"), 
                                names = stringr::str_remove(names, pattern = "X"), 
                                DOY = lubridate::as_date(names),
                week = lubridate::week(DOY))

# south region

"names" <- NULL
"values" <- NULL
df<- data.frame()
for(i in 1:raster::nlayers(south_uwind)){
  names <- south_uwind[[i]]@data@names
  mean_wind <- mean(south_uwind[[i]]@data@values, na.rm = TRUE)
  min_wind <- south_uwind[[i]]@data@min
  max_wind <- south_uwind[[i]]@data@max
  
  dat<-cbind(names, mean_wind, min_wind, max_wind)
  df<-rbind(df, dat)
  }
head(df)

south_df <- df %>%
  dplyr::mutate(Year = stringr::str_extract(names, pattern = "\\d{4}"), 
                                names = stringr::str_remove(names, pattern = "X"), 
                                DOY = lubridate::as_date(names),
                week = lubridate::week(DOY))

# plots

monthly_wind <- rbind(north_df %>% 
                      dplyr::mutate(region = "North"),
                    south_df %>% 
                      dplyr::mutate(region = "South")) %>%
  dplyr::mutate(month = lubridate::month(DOY),
                year = lubridate::year(DOY))

u_monthly_wind <- monthly_wind

ggplot2::ggplot(monthly_wind,
                ggplot2::aes(x = month,
                             y = mean_wind %>%
                               as.numeric(),
                            # shape = region,
                             color = as.factor(year)
                             )) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme_bw() +
  ggplot2::facet_wrap(~region,
                      ncol = 1) +
  #viridis::scale_color_viridis() +
  ggplot2::ggtitle("Monthly u wind") +
  ggplot2::theme(legend.position = "bottom")
```

##### Test correlation with recruitment index from the 2019 assessment model
```{r, fig.height = 16}
recruit <- NEesp::asmt %>% 
  dplyr::filter(Species == "Bluefish", 
                Metric == "Recruitment", AssessmentYear == 2019) %>%
  dplyr::select(Year, Value) %>%
  dplyr::rename(Recruitment = Value) %>%
  dplyr::mutate(Year = as.character(Year))

dplyr::full_join(recruit, monthly_wind) %>%
  tidyr::drop_na(Recruitment, mean_wind) %>%
  ggplot2::ggplot(ggplot2::aes(x = as.numeric(mean_wind),
                               y = Recruitment)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::facet_grid(cols = ggplot2::vars(region),
                      rows = ggplot2::vars(month)) +
  ggplot2::theme_bw()
```


```{r}
dat <- dplyr::full_join(v_monthly_wind %>%
                   dplyr::rename(v_mean_wind = mean_wind) %>%
                   dplyr::select(-c(min_wind, max_wind)),
      u_monthly_wind %>%
                   dplyr::rename(u_mean_wind = mean_wind) %>%
                   dplyr::select(-c(min_wind, max_wind))) %>%
  dplyr::full_join(recruit) %>%
  tidyr::drop_na()

dat %>%
  ggplot2::ggplot(ggplot2::aes(x = as.numeric(u_mean_wind),
                               y = as.numeric(v_mean_wind),
                               color = log(Recruitment))) +
  ggplot2::geom_point(cex = 2) +
  ggplot2::theme_bw() +
  ggplot2::facet_grid(rows = ggplot2::vars(month),
                      cols = ggplot2::vars(region)) +
  viridis::scale_color_viridis()
```