---
title: "Bluefish temperature analysis"
author: "Abigail Tyrell"
date: "`r format(Sys.time(), '%d %b %Y')`"
output:
  html_document:
    toc: true
    number_sections: true
    toc_depth: 4
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
`%>%` <- magrittr::`%>%`
```

# Introduction

This is a proof-of-concept to illustrate the calculation of a temperature indicator for the bluefish ESP: location of the 22C isotherm in July.

# Shapefile setup
```{r}
xmin <- -80
xmax <- -65
ymin <- 30
ymax <- 45
xlims <- c(xmin, xmax)
ylims <- c(ymin, ymax)
crs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

ne_countries <- rnaturalearth::ne_countries(
  scale = 10,
  continent = "North America",
  returnclass = "sf"
) %>%
  sf::st_transform()

ne_states <- rnaturalearth::ne_states(
  country = "united states of america",
  returnclass = "sf"
) %>% sf::st_transform()

new_shape <- NEesp::shape %>%
  dplyr::select(STRATA, geometry) %>%
  sf::st_transform(proj4string = crs)

# two masks
msk <- ecodata::epu_sf %>%
  dplyr::filter(EPU == "MAB") %>%
  dplyr::select(geometry) %>%
  sf::st_transform(proj4string = crs)

large_geom <- new_shape %>% 
    dplyr::summarise(geometry = sf::st_union(geometry)) %>%
  sf::st_crop(y = c(xmin = -80, xmax = -69, 
                    ymax = 41.5, ymin = 32))

# create mab without inshore and sab by breaking large_geom
mab <- large_geom %>%  
  sf::st_crop(y = c(xmin = -80, xmax = -69, 
                    ymax = 41.5, ymin =  35.8327))

sab <- large_geom %>%  
  sf::st_crop(y = c(xmin = -80, xmax = -69, 
                    ymax =  35.8327, ymin = 32))

# base plot
p1 <- ggplot2::ggplot() +
  ggplot2::geom_sf(
    data = new_shape, 
    alpha = 0.9, 
    size = 0.01, 
    color = "grey30",
    fill = "white"
  ) +
  ggplot2::geom_sf(
    data = ne_countries,
    color = "grey60", 
    size = 0.25
  ) +
  ggplot2::geom_sf(
    data = ne_states,
    color = "grey60", 
    size = 0.05
  ) +
  viridis::scale_fill_viridis(discrete = FALSE,
                              na.value = NA) +
  ggplot2::coord_sf(crs = crs, 
                    xlim = xlims, 
                    ylim = ylims) +
  ggthemes::theme_map() +
  ggplot2::theme(legend.direction = "horizontal")
```

# 2021

## Download data

For some reason, this doesn't work in desktop R/RStudio, but it does work in github actions. Not evaluating this chunk for now and using a manually downloaded file instead. The whole workflow can be put into github actions as a workaround.
```{r, eval = FALSE}
  name <- "2022.nc"
  dir.create(here::here("data-raw","gridded", "sst_data"), recursive = TRUE)
  filename <- here::here("data-raw","gridded", "sst_data", "test_2022.grd")
  url <- paste0("https://downloads.psl.noaa.gov/Datasets/noaa.oisst.v2.highres/sst.day.mean.", "2022", ".v2.nc")
  download.file(url, destfile = name)
```

## Convert to raster
```{r}
test_2021 <- ecopull::nc_to_raster(nc = here::here("sst.day.mean.2021.v2.nc"), varname = 'sst')
test_2021 <- raster::rotate(test_2021)
```

## Crop to MAB (no inshore)
```{r}
mab_temp <- raster::mask(x = test_2021[[1:180]], 
                        mask = mab)
mab_temp2 <- raster::mask(x = test_2021[[181:365]], 
                        mask = mab)

raster::plot(mab_temp[[1]])
```

### Plot over map
```{r}
rast_mab_df <- raster::as.data.frame(mab_temp, xy = TRUE)
rast_mab_df2 <- raster::as.data.frame(mab_temp2, xy = TRUE)

p1 +
  ggplot2::geom_tile(data = rast_mab_df[,1:3] %>%
                       tidyr::drop_na(),
                     ggplot2::aes(x = x,
                                  y = y,
                                  fill = X2021.01.01)) +
  ggplot2::theme(legend.direction = "horizontal")
```

#### July 1
```{r}
p1 +
  ggplot2::geom_tile(data = rast_mab_df2,
                     ggplot2::aes(x = x,
                                  y = y,
                                  fill = X2021.07.01)) +
  ggplot2::theme(legend.direction = "horizontal")
```

#### July 31
```{r}
p1 +
  ggplot2::geom_tile(data = rast_mab_df2,
                     ggplot2::aes(x = x,
                                  y = y,
                                  fill = X2021.07.31)) +
  ggplot2::theme(legend.direction = "horizontal")
```

## Calculate proportion of area above 22C in July by day
```{r}
## mab

names <- c()
value <- c()
for(i in 3:ncol(rast_mab_df)){
  new_dat <- rast_mab_df[,i] %>%
    tibble::as_tibble() %>%
    tidyr::drop_na()
  names[i-2] <- colnames(rast_mab_df[i])
  value[i-2] <- nrow(new_dat %>% dplyr::filter(value > 22)) / nrow(new_dat)
}
mab_prop <- tibble::tibble(names, value, region = "MAB")

names <- c()
value <- c()
for(i in 3:ncol(rast_mab_df2)){
  new_dat <- rast_mab_df2[,i] %>%
    tibble::as_tibble() %>%
    tidyr::drop_na()
  names[i-2] <- colnames(rast_mab_df2[i])
  value[i-2] <- nrow(new_dat %>% dplyr::filter(value > 22)) / nrow(new_dat)
}
mab_prop2 <- tibble::tibble(names, value, region = "MAB")

mab_prop <- rbind(mab_prop, mab_prop2)

## combine
mab_prop <- mab_prop %>%
  dplyr::mutate(Year = stringr::str_extract(names, pattern = "\\d{4}"), 
                                names = stringr::str_remove(names, pattern = "X"), 
                                DOY = lubridate::as_date(names),
                week = lubridate::week(DOY),
                month = lubridate::month(DOY),
                value = as.numeric(value))

## plot
ggplot2::ggplot(mab_prop %>%
                  dplyr::filter(month == 7),
                ggplot2::aes(x = DOY,
                             y = value,
                             color = region)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme_bw() +
  ggplot2::ggtitle("Daily proportion of cells above 22C in July") +
  ggplot2::theme(legend.position = "none")
```

### Mean proportion of area over 22C in July
```{r}
mab_prop %>%
  dplyr::filter(month == 7) %>%
  dplyr::summarise(mean = mean(value) %>% round(digits = 3),
                   sd = sd(value) %>% round(digits = 3))
```